Unsupervised embedding of trajectories captures the latent structure of mobility

### 2021-03-10 paper/rebuttal/rebuttal.tex

\makeatletter
\renewcommand{\maketitle}{\bgroup\setlength{\parindent}{0pt}
\begin{flushleft}
\Large  \textbf{\@title}
\end{flushleft}\egroup
}
\makeatother
\title{Re: Nature submission 2020-10-19515}
\date{}

\newgeometry{margin=1in, top=1cm}
\begin{document}
\maketitle

\bigskip
\thispagestyle{empty}
\vspace{-0.4cm}
\noindent Dear Dr. Sutherland and Referees,\\
\vspace{-0.25cm}

We thank all the referees for their thoughtful and helpful comments and critiques.
We acknowledge that our initial submission had a substantial gap regarding the choice of the \textit{word2vec} model, its connection with the gravity model of mobility, and its evaluation.
While reflecting on these comments and our manuscript, we have made a theoretical breakthrough that addresses the fundamental issues, and dramatically expands the scope of our contribution;
namely, we have discovered \emph{a mathematical equivalence between the \textit{word2vec} (skip-gram with negative sampling) model and the gravity model.}

@@

Furthermore, our subsequent experiments show that our approach's performance is robust, exceeding many other baselines, even direct optimization of the gravity model.
Therefore, with these changes, we are confident that our paper is substantially stronger with a qualitative leap.
Thus, we would like to request a reconsideration of our manuscript.
Again, we are incredibly appreciative of these comments, which provided the insight and motivation needed to make these breakthroughs.

We have developed the mathematical and theoretical basis of our approach for embedding mobility, and made additional updates to our experiments and manuscript, addressing the three main issues raised by referees:
\vspace{-0.1cm}
\begin{enumerate}
	\itemsep0em
	\item The lack of theoretical justification and grounding for the application of \textit{word2vec}
	\item The lack of simpler baselines to understand the performance of \textit{word2vec}
	\item The lack of sufficient grounding in the literature of neural embeddings and mobility
\end{enumerate}
\vspace{-0.1cm}
We address the first point by presenting the mathematical equivalence between the Gravity Model of Mobility and the skip-gram negative sampling \textit{word2vec} model.
Namely, we demonstrate that \textit{word2vec}'s objective function is equivalent to the gravity model and thus \textit{word2vec} learns gravity-like relationships between entities, whether it be language or locations.
This development constitutes a major breakthrough in our understanding of neural embeddings, and in addition to justifying their use in our study, also opens the door for further research applying them to the study of mobility.

@@

	Notably, the location embedding methodology in Liu et al (2016) (see Section 3.1) is essentially the same as the approach used in the submitted manuscript, and the work by Liu et al (2016) is well-known with over 100 citations.
}

\response{Thank you for pointing out the missing references!
	We have carefully read through each, and learned a lot in the process.
	They have now been incorporated into our manuscript's introduction.
	Although there are existing works on trajectory-based mobility embedding, we argue that our new result establishing the mathematical equivalence of \textit{word2vec} and the gravity model, as well as our extensive evaluation with many baselines represents a novel and substantial contribution.
}

\rcomment{
	I found the specific use of the SGNS algorithm unmotivated. There are many ways to generate useful low-dimensional embeddings, so why is the SGNS method prioritized above others? This method is popular, but popularity alone is not sufficient justification and I see no theoretical reason why it should be the most appropriate method. For example, a key point in this manuscript is the fact that the the SGNS embeddings capture the flux between locations better than geographic distance. However, if this is such an important point, then why not simply learn a low-dimensional embedding approximation of the flux metric directly? such as SVD?
}

\response{
	Thank you for the suggestion!
	Since our original submission, we have uncovered a mathematical connection between the skip-gram model with negative sampling, and the gravity model of mobility.
	Namely, we find that \emph{SGNS is a gravity model}.
	This finding explains why SGNS performs so well and demonstrates the utility of the model for studying contexts where gravity relationships are expected, such as in human mobility.
	We refer you to Section 1 of this letter, above, for an investigation of this connection, which we believe to be not only a justification of our choice of the \textit{word2vec} model to study mobility, but also a major contribution to the understanding of neural embeddings and the gravity model themselves.

### 2021-03-10 paper/current/main.tex
add some text emphasizing contribution

	Human mobility and migration drive major societal phenomena such as the growth and evolution of cities~\autocite{wef2017migration, curiel2018citygravity},  epidemics~\autocite{kraemer2020covid, truscott2012epidemicgravity}, economies~\autocite{kaluza2010cargo, kerr2011immigration}, and innovation~\autocite{kaiser2018innovation, sugimoto2017mostimpact, petersen2018multiscale}.
	Historically, human mobility has been strongly constrained by physical separation---geographic distance~\autocite{zipf1946gravity}.
	However, geographic distance is becoming less relevant in the increasingly-globalized world~\autocite{czaika2018globalisation, buchan2009globalization} in which physical barriers are shrinking while linguistic, cultural, and historical relationships are becoming more important.
	As understanding mobility is becoming critical for contemporary society, finding frameworks that can capture this complexity is of paramount importance.
	Here, we reveal a mathematical connection between the neural embedding model \textit{word2ec}~\autocite{mikolov2013word2vec} and the gravity model of mobility and, using three distinct human trajectory datasets, demonstrate that a neural embedding model can encode nuanced relationships between locations into a vector-space, providing an effective measure of distance that reflects the multi-faceted structure of human mobility.
	Focusing on the case of scientific mobility, we show that embeddings of scientific organizations uncover cultural and linguistic relations, and even academic prestige, at multiple levels of granularity.
	Furthermore, the embedding vectors reveal universal relationships between organizational characteristics and their place in the global landscape of scientific mobility.
	The ability to learn scalable, dense, and meaningful representations of mobility directly from the data can open up a new avenue of studying mobility across domains.

@@

While we focus on three domains of mobility, this approach may be broadly applicable to other kinds of mobility data that is expected to follow a gravity-like pattern, such as general human migration, transit-network mobility, and more.
Moreover, this approach can be used to learn a functional distance even between entities for which no physical analog exists, such as between occupational categories based on individuals' career trajectories.
In addition to providing a functional distance that supports modeling and predicting mobility patterns, the structure of the neural embedding space is amenable to a range of unique applications for studying mobility.
As we have shown, the embedding space allows the visualization of the complex structure of scientific mobility at high resolution across multiple scales, providing a large and detailed map of the landscape of global scientific mobility.
Embedding also allows us to quantitatively explore abstract notions such as academic prestige, and can potentially be generalized to other abstract axes.
Investigation of the structure of the embedding space, such as the vector norm, reveals universal patterns based on the organization's size and their vector norm that could be leveraged in future studies of mobility.

In spite of its promise, our study has several limitations.
First, the skip-gram \textit{word2vec} model does not leverage directionality, meaning that embedding will be less effective at capturing mobility for which directionality is critical.
Future studies may also consider bi-directional embeddings, such as BERT~\autocite{devlin2018bert}, to incorporate directionality.
Second, the neural embedding approach is most useful in cases of mobility between discrete geographic units such as between countries, cities, and businesses; this approach is less useful in the case of mobility between locations represented using geographic coordinates, such as in the modeling of animal movements.
Neither of these methodological limitations is insurmountable, and future work can aim to incorporate directionality and identify meaningful representations of continuous mobility data.
Third, neural embeddings are an inherently stochastic procedure, and so results may change across different iterations.
We note that all results shown here are robust to this stochasticity, likely emerging from the limited ``vocabulary'' of scientific mobility, airports, and accommodations (several thousand) and the massive datasets used (several million trajectories).
Applications of \textit{word2vec }to problem domains where the ratio of the vocabulary to data is smaller, however, should take caution to ensure that findings are not the result of random fluctuations.
Finally, the case of scientific mobility presents domain-specific limitations.
Reliance on bibliometric metadata means that we capture only long-term mobility such as migration, rather than the array or more frequent short-term mobility such as conference travel and temporary visits.
The kinds of mobility we do capture---migration, and co-affiliation---although conceptually different, are treated identically by our model.
Also, our data might further suffer from bias based on publication rates: researchers at prestigious organizations tend to have more publications, leading to these organizations appearing more frequently in affiliation trajectories.

Mobility and migration are at the core of human nature and history, driving societal phenomena as diverse as epidemics\autocite{kraemer2020covid, truscott2012epidemicgravity} and innovation~\autocite{kaiser2018innovation, sugimoto2017mostimpact, petersen2018multiscale, morgan2018prestige, rodrigues2016mobility}.
However, the paradigm of scientific migration may be changing.
Traditional hubs of migration have experienced many politically-motivated policy changes that affect scientific mobility, such as travel restrictions in the U.S. and U.K.~\autocite{chinchilla2018travelban}.
At the same time, other nations, such as China, are growing into major scientific powers and attractors of talent~\autocite{cao2020returning}.
Unprecedented health crises such as the COVID-19 pandemic threaten to bring drastic global changes to travel and migration by tightening borders and halting travel.
With the changing paradigm of global mobility, we need now, more than ever, new tools and approaches to capture and understand human mobility in order to inform sensible, effective, sustainable, and humane policies.

>>>>

	Human mobility and migration drive major societal phenomena such as the growth and evolution of cities~\autocite{wef2017migration, curiel2018citygravity},  epidemics~\autocite{kraemer2020covid, truscott2012epidemicgravity}, economies~\autocite{kaluza2010cargo, kerr2011immigration}, and innovation~\autocite{kaiser2018innovation, sugimoto2017mostimpact, petersen2018multiscale}.
	Historically, human mobility has been strongly constrained by physical separation---geographic distance~\autocite{zipf1946gravity}.
	However, geographic distance is becoming less relevant in the increasingly-globalized world~\autocite{czaika2018globalisation, buchan2009globalization} in which physical barriers are shrinking while linguistic, cultural, and historical relationships are becoming more important.
	As understanding mobility is becoming critical for contemporary society, finding frameworks that can capture this complexity is of paramount importance.
	Here, we reveal that the neural embedding model \textit{word2ec}~\autocite{mikolov2013word2vec} is mathematically equivalent to the gravity model of mobility and, using three distinct human trajectory datasets, demonstrate that a neural embedding model can encode nuanced relationships between locations into a vector-space, providing an effective measure of distance that reflects the multi-faceted structure of human mobility.
	Focusing on the case of scientific mobility, we show that embeddings of scientific organizations uncover cultural and linguistic relations, and even academic prestige, at multiple levels of granularity.
	Furthermore, the embedding vectors reveal universal relationships between organizational characteristics and their place in the global landscape of scientific mobility.	
	The ability to learn scalable, dense, and meaningful representations of mobility directly from the data can open up new avenues for the study of mobility across domains.

@@

In revealing the correspondence between neural embeddings and the gravity model, the study of human mobility can move beyond geographic and network-based models of mobility, and instead leverage the high-order structure from individuals' mobility trajectories. 
While we focus on three domains of mobility, this approach could be applied to many different domains, such as general human migration, transit-network mobility, and more.
Once learned, functional distances between locations, such as countries, cities, or organizations, can be published to support future research.
Moreover, this approach can be used to learn a functional distance even between entities for which no geographic analog exists, such as between occupational categories based on individuals' career trajectories. 
In addition to providing a functional distance that supports modeling and predicting mobility patterns, the structure of the neural embedding space is amenable to a range of unique applications for studying mobility.
As we have shown, the embedding space allows the visualization of the complex structure of scientific mobility at high resolution across multiple scales, providing a large and detailed map of the landscape of global scientific mobility.
Embedding also allows us to quantitatively explore abstract notions such as academic prestige, and can potentially be generalized to other abstract axes.
Investigation of the structure of the embedding space, such as the vector norm, reveals universal patterns based on the organization's size and their vector norm that could be leveraged in future studies of mobility.

This approach, and our study, also have several limitations.
First, the skip-gram \textit{word2vec} model does not leverage directionality, meaning that embedding will be less effective at capturing mobility for which directionality is critical.
Future studies may consider bi-directional embeddings, such as BERT~\autocite{devlin2018bert}, to incorporate directionality, as well as their correspondance to asymetric mobility models, such as the radiation model~\cite{simini2012universal}.
Second, the neural embedding approach is most useful in cases of mobility between discrete geographic units such as between countries, cities, and businesses; this approach is less useful in the case of mobility between locations represented using geographic coordinates, such as in the modeling of animal movements.
Third, neural embeddings are an inherently stochastic procedure, and so results may change across different iterations.
However, in this study we observe all results to be robust to stochasticity, likely emerging from the limited ``vocabulary'' of scientific mobility, airports, and accommodations (several thousand) and the relatively0massive datasets used to learn representations (several million trajectories).
Applications of \textit{word2vec} to problem domains where the ratio of the vocabulary to data is smaller, however, should take caution to ensure that findings are not the result of random fluctuations.
Finally, the case of scientific mobility presents domain-specific limitations.
Reliance on bibliometric metadata means that we capture only long-term mobility such as migration, rather than the array or more frequent short-term mobility such as conference travel and temporary visits.
The kinds of mobility we do capture---migration and co-affiliation---although conceptually different, are treated identically by our model.
Also, our data might further suffer from bias based on publication rates: researchers at prestigious organizations tend to have more publications, leading to these organizations appearing more frequently in affiliation trajectories.

Mobility and migration are at the core of human nature and history, driving societal phenomena as diverse as epidemics\autocite{kraemer2020covid, truscott2012epidemicgravity} and innovation~\autocite{kaiser2018innovation, sugimoto2017mostimpact, petersen2018multiscale, morgan2018prestige, rodrigues2016mobility}.
However, the paradigm of scientific migration may be changing.
Traditional hubs of migration have experienced many politically-motivated policy changes that affect scientific mobility, such as travel restrictions in the U.S. and U.K.~\autocite{chinchilla2018travelban}.
At the same time, other nations, such as China, are growing into major scientific powers and attractors of talent~\autocite{cao2020returning}.
Unprecedented health crises such as the COVID-19 pandemic threaten to bring drastic global changes to travel and migration by tightening borders and halting travel.
By revealing the correspondence between neural embedding and the gravity model and revealing their utility and efficacy, our study opens a new avenue in the study of mobility.
Mobility is at the core of many global challenges, and the insights brought forth by our tool can be put to use to inform a better understanding of human mobility, and to support inform a sensible, effective, sustainable, and humane policies.
